---
title: 'Examination Project'
author: 'Christina Autoshi Baidya'
email: 'Baidya.Christinaautoshi.7374@student.uu.se'
date: '04-12-18'
mainfont: 
sansfont: 
mathfont: 
monofont: 
fontsize: 
output: 
 html_document: 
    code_folding: show
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: false
      smooth_scroll: true
---

```{r load packages, message = FALSE}
library(tidyr)    # used for putting data into long format
library(knitr)    # used for creation of the report

```



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy.opts = list(width.cutoff = 60))
```




# Environment setup

```{r define drug concentrations path}
dir_name_drug  <- '../desktop/exam project'
file_name_drug  <- 'Big_Pharma_BPI889_data_200mg_26.csv'
file_path_drug  <- paste(dir_name_drug, file_name_drug, sep = '')    #These commands for drug file directory 
```
```{r define mutation concentrations path}
dir_name_mutation  <- '../desktop/exam project'
file_name_mutation  <- 'stu26.csv'
file_path_mutation <- paste(dir_name_mutation, file_name_mutation, sep = '')    #These commands for mutation file directory
```


# data import

```{r read-in drug}
read.csv(file = 'Big_Pharma_BPI889_data_200mg_26.csv', header = TRUE, as.is = TRUE)
data_pk <- read.csv(file = 'Big_Pharma_BPI889_data_200mg_26.csv', header = TRUE, as.is = TRUE, na.strings = ',')  #read.csv command to read csv file
```


```{r read-in mutation}
read.csv(file = 'stu26.csv', header = TRUE, as.is = TRUE)
data_snp <- read.csv(file = 'stu26.csv', header = TRUE, as.is = TRUE, na.strings = ',') #read.csv command to read csv file
```

# Data check-out
This section should contain any data check-out.

### Structure of the  data sets

```{r drug structure}
str(data_pk)
```


```{r drug head}
head(data_pk)
```


```{r mutation structure}
str(data_snp)
```

```{r mutation head}
head(data_snp)
```


```{r drug}
colnames(data_pk)
```
```{r mutation}
colnames(data_snp)
```

### Head of the `drug` Data Set


```{r knitr kable example, eval = FALSE}
kable(x = data_pk,
      align = 'c',
      row.names = FALSE,
      caption = 'individual data')
```


```{r drug head table}
head_PK <- head(data_pk, n = 5L)

```


### Head of the `mutation` Data Set

```{r knitr kable example, eval = FALSE}
kable(x = data_snp,
      align = 'c',
      row.names = FALSE,
      caption = 'individual data')
```

```{r mutation head table}
head_SNP <- head(data_snp, n = 5L)

```

# Data management 


This section should contain all data manipulation needed to perform the examination project.

## Set `ID` Column Name

```{r mutation id column}
colnames(data_snp)[1] <- c('ID')  #change the column name `X` to `ID`
```

```{r drug id column}
colnames(data_pk)[1] <- c('ID')   #change the column name `X` to `ID`
```


###Convert mutation data as widtype, heterozygote, homozygote factor

```{r Create new data}
data_snp2 <- data_snp       #Create duplicate data of mutation
```

```{r convert T134A to factor}
data_snp2[, c("T134A","A443G","G769C","G955C","A990C","G1051A","G1078T","T1941A","T2138C","G2638T")] <- lapply(data_snp2[, 
            c("T134A","A443G","G769C","G955C","A990C","G1051A","G1078T","T1941A","T2138C","G2638T")],
                      levels = c('0', '1', '2'),
                      labels = c('widtype', 'heterozygote','homozygote'),factor)
```


```{r tidy mutation data}
data_snp2 <- gather(data  = data_snp2,
                  key   = Mutation,
                  value = Type, -ID)    #Gather all data of `data_snp2` unless `ID` column
```

```{r}
kable(x = data_snp2,
      align = 'c',
      row.names = FALSE,
      caption = 'individual data')
```


###Set `Time` column name

```{r Create new Data}
data_pk2 <- data_pk     #Create duplicate data of drug
```


```{r drug Time column}
colnames(data_pk2)[2] <- c('Time.30m')
colnames(data_pk2)[4] <- c('Time.2h')
colnames(data_pk2)[6] <- c('Time.6h')
colnames(data_pk2)[7] <- c('Time.12h')
colnames(data_pk2)[8] <- c('Time.24h')      #Set all Column names same as `Time.`
```

### Transform the Data Set to Long Format

```{r tidy PK_02 data}
data_pk2 <- gather(data  = data_pk2,
                  key   = TIME,
                  value = Concen, -ID , -Sex, -Weight..kg., -Height..cm.)    #Gather all data of `data_snp2` unless column `ID`,`Sex`,`Weight..kg.`,`Height..cm.`
```

### Create Numeric `Concentration` Column
```{r}
data_pk2$Concen = as.numeric(data_pk2$Concen)
```


```{r PK_02 head long}
nrow(data_pk2)
ncol(data_pk2)
head_PK2 <- head(data_pk2, n = 5L)
kable(x = head_PK2,
      align = 'c',
      row.names = FALSE,
      caption = 'individual data-long format')
```

### Remove `Time` from time
```{r convert time numeric}
data_pk2$TIME <- gsub('Time.','',data_pk2$TIME)   #Remove `Time.` from column `TIME`
```

```{r}
kable(x = data_pk2,
      align = 'c',
      row.names = FALSE,
      caption = 'individual data')
```

###Save data as `rdata` formate
```{r save cov data set as RData}
dir <- '../new R/'
save(data_pk2, file = 'Big_Pharma_BPI889_data_200mg_26.rdata')
```



# Data summary
This section should contain the numerical data exploration for the project team.

##summary full data
```{r summary all data}
data_merge <- merge(data_pk2, data_snp2)
summary(data_merge)
```


### Create Numeric `Weight..kg. & Weight..kg.` Column
```{r convert Weight..kg. & Height..cm. numeric}
data_merge$Weight..kg. = as.numeric(data_merge$Weight..kg.)
data_merge$Height..cm. = as.numeric(data_merge$Height..cm.)
```


##Summary individually
```{r merged data}
summary(data_merge[,c('Weight..kg.', 'Height..cm.','Concen')])   #Summary of Weight..kg, Height..cm, Concen
```



```{r merged data}
summary(data_merge[,c('Mutation','Type')])    #Summary of Mutation & Type
```


#Create Unique data
```{r  merged data}
data_unique <- data_merge[!duplicated(data_merge$ID), ]
```


# Graphical exploration
This section should contain the graphical data exploration for the project team.

### Distribution
```{r histograms, fig.height = 5, fig.width = 5}
par(mfrow = c(1, 3))
for (cov in c('Weight..kg.', 'Height..cm.', 'Concen')) { 
  # Add code for the plot here
  hist(x = data_unique[,cov], xlab = NA, main = cov)
  }
```

### Correlation Matrix
```{r pairs plot}
# This is an example do not write code here
pairs(data_unique[,c('Weight..kg.', 'Height..cm.', 'Concen')],
      upper.panel = NULL,
      lower.panel = panel.smooth, 
      pch = 20)
```

### Box Plots
```{r boxplots}
par(mfrow = c(2, 2))
boxplot(Weight..kg.~Height..cm., data = data_unique, main = 'Weight..kg.~Height..cm.')
boxplot(Height..cm.~Concen, data = data_unique, main = 'Height..cm.~Concen')
boxplot(Weight..kg.~Concen, data = data_unique, main = 'Weight..kg.~Concen')

```

# Pharmacokinetic Calculations
This section should contain relevant calculations of PK parameters.
#Create new data frame
```{r}
data_pk3 <- data_pk     #Create duplicate data of drug
```

```{r}
data_pk3$Time...12h = as.numeric(data_pk3$Time...12h)
data_pk3$Time...24h = as.numeric(data_pk3$Time...24h)
```


##Find the value of K
```{r find the value of K}
data_pk3[c(8,10,14),'K'] <- (-(log(data_pk3[c(8,10,14),'Time...12h']) - log(data_pk3[c(8,10,14),'time.6h'])) / (12-6))
data_pk3$K <- round(data_pk3$K, 3)

```

##Find the missing value of Time...24
```{r find the missing value}
data_pk3[c(8,10,14), 'Time...24h'] <- round((data_pk3[c(8,10,14),'Time...12h']) * exp(-(data_pk3[c(8,10,14),'K'])*(24-12)),3)  #caltulation and roun together
```

##Find the value of AUC
```{r find the missing value}
data_pk3$AUC <-  round(((data_pk3$Time...24h) + (data_pk3$Time...12h)) * (1/5) * (24-12),3)  #caltulation and roun together
```

##Find the value of CL
```{r find CL}
data_pk3$CL <- round(200 / data_pk3$AUC, 1) 
```

```{r}
data_pk3$Cav <-  round ((200 * 0.9) / (data_pk3$CL * data_pk3$Time...24h),3)
```

```{r}
kable(x = data_pk3,
      align = 'c',
      row.names = FALSE,
      caption = 'individual data')
```

# Covariate testing of PK parameters
This section should contain relevant statistical testing of PK parameters and subject demographics.

# Assessment of clinical impact
This section should contain relevant evaluation of the clinical impact of the relationships identified in section above.

